---
title: "Learn_tidyverse_01"
author: "tmyst"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: zenburn
    md_extension: -ascii_identifiers
    self_contained: true
    lightobx: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
```

```{r include=FALSE}
path <- "/Users/tomoya/Rprojects/GitHubShare/DSNotebooks/Data/datasets_99214_241651_mr-boston-flattened.csv"
cocktails <- read_csv(path)
```


Tidy data is data where:
1. Every column is variable.
2. Every row is an observation.
3. Every cell is a single value.

```{r}
cocktails %>% count(category, sort = T)
cocktails
```

tidyr functions fall into five main categories:

1. Pivotting
2. Rectangling
3. Nesting
4. Splitting and combining of character values
5. Deatling with NA

```{r}
cocktails %>% count(name) %>% filter(n>1) # confirm that name can be the key
cocktails %>% 
  pivot_longer(cols = starts_with("measurement"), 
               names_to = "measurement_index", 
               values_to = "measurement_result"
  )
```
But, we have other 6 columns containing ingredients.
```{r}
cocktails_tidy <- cocktails %>% 
  pivot_longer(cols = c(starts_with("ingredient"), starts_with("measurement")),
    names_to = c("value_type", "index"), 
    names_sep = "-"
  ) %>% 
  pivot_wider(
    names_from = value_type,
    values_from = value
  )
cocktails_tidy
```

Or, if we'd like to examine how to reshape the data.frame in pivot_longer operation,
we can use "spec" objects. "spec" can be manually modified
```{r}
spec <- cocktails %>% 
  build_longer_spec(
    cols = c(starts_with("ingredient"), starts_with("measurement")),
    names_to = c("value_type", "index"), 
    names_sep = "-"
  )
spec2 <- spec %>% mutate(.value = value_type) %>% select(-value_type)

list(
  spec2,
  cocktails %>% pivot_longer_spec(spec2) 
)
```

Or we can overwrite column ".value". The simplest operation will be like the following,
```{r}
spec <- cocktails %>% 
  build_longer_spec(
    cols = c(starts_with("ingredient"), starts_with("measurement")),
    names_to = c(".value", "index"), 
    names_sep = "-"
  )

list(
  spec,
  cocktails %>% pivot_longer_spec(spec)
)
```

```{r}
cocktails_tidy %>% count(name)
```

